---
title: "**Analysis**"
output: 
  html_document:
      toc: true
      toc_float: true
      code_folding: hide
---

lost variables: ecigs_past_month, ads_ecigs, harm_ecigs, age_first_ecig, reasons_ecigs

```{r}
library(tidyverse)
library(srvyr)
```




```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "magma",
  ggplot2.continuous.fill = "magma"
)

scale_colour_discrete = scale_color_viridis_d(option = "magma")
scale_fill_discrete = scale_fill_viridis_d(option = "magma")

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r}
ecig_data = read.csv("./Data/ecig_data.csv")  
  
  
```



# Logistic Model

tbd





# Ad Viewing

```{r, echo = F, message=F}
ecig_data %>% 
  mutate(
    ecigs_past_month = replace(ecigs_past_month, ecigs_past_month == "1-2", "1.5"),
    ecigs_past_month = replace(ecigs_past_month, ecigs_past_month == "3-5", "4"),
    ecigs_past_month = replace(ecigs_past_month, ecigs_past_month == "6-9", "7.5"),
    ecigs_past_month = replace(ecigs_past_month, ecigs_past_month == "10-19", "14.5"),
    ecigs_past_month = replace(ecigs_past_month, ecigs_past_month == "20-29", "24.5"),
                                    
    ecigs_past_month = as.numeric(ecigs_past_month), 
    ads_ecigs = replace(ads_ecigs, ads_ecigs == "do no use the internet", "do not use the internet"), 
    ads_ecigs = fct_relevel(ads_ecigs, c("never", "rarely", "sometimes", "most of the time", "always", "do not use the internet"))
  ) %>% 
  select(ecigs_past_month, ads_ecigs, weight) %>% 
  drop_na() %>% 
  filter(ads_ecigs != "do not use the internet") %>% 
  group_by(ads_ecigs) %>% 
  summarize('Avg days Consuing of E-Cig Use' = round(weighted.mean(ecigs_past_month, as.numeric(weight)), digits = 3)) %>% 
  rename("Viewing Rate of Internet E-Cig Ads" = ads_ecigs) %>%
  knitr::kable(digits = 2)
```


# Timeline 

Does electronic cigarette use differ across year and sex?
```{r, echo = F, message=F}
ecig_data %>%
  filter(is.na(try_ecigs) == FALSE, is.na(sex) == FALSE) %>%
  group_by(year, sex) %>%
  mutate(indc = ifelse(try_ecigs == "yes", 1, 0)) %>%
  summarise(y = weighted.mean(indc, as.numeric(weight))) %>%
  ggplot(aes(x = year, y = y, fill = sex)) +
  geom_col(position = "dodge") +
  #geom_text(aes(label=len), vjust=1.6, color="white", size=3.5) +
  ylab("Proportion Tried E-cigarettes") +
  xlab("Year") +
  scale_fill_manual(values = c("#75CAF0", "#111ED0")) 
```



Has the perception of harmfulness of electronic cigarettes compared to cigarettes changed over time?
```{r, echo = F, message=F}
#creating survey object 
library(srvyr)
library(viridis)
srv_ecig = ecig_data %>% 
  mutate(id = row_number()) %>% 
   as_survey_design(id = id, wt = as.numeric(weight) )

#plot of harmful vs year 
srv_ecig %>% 
  filter( !is.na(quit_cig) , quit_cig != "not a smoker" ) %>% 
  filter( !is.na(harm_ecigs) ) %>%
  mutate(harm_ecigs = recode(harm_ecigs, "equally addictive" = "equally", "less addictive" = "less", "more addictive" = "more"), 
         ecigs_past_month = factor(ecigs_past_month, c(NA, "0", "1-2", "3-5", "6-9", "10-19", "20-29", "30"))
         ) %>% 
  filter( harm_ecigs =="less"| harm_ecigs =="equally" | harm_ecigs =="more") %>%
  filter( year != "2013", year != "2015") %>% 
  mutate(harm_ecigs = fct_relevel(harm_ecigs, c("less", "equally", "more"))) %>% 
  group_by(year, harm_ecigs) %>% 
  summarize(n = survey_total()) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes( x = year, y = prop, fill = harm_ecigs)) + 
  geom_bar(position ="dodge", stat ="identity") +
  scale_fill_viridis(discrete = T) +
  labs(
    title = "Perception of Harmfulness of E-Cigarettes Compared to Cigarettes",
    x = "Year",
    y = "Proportion of Belief") +
  theme(plot.title = element_text(hjust = 0.5))
```

Has the frequency of electronic cigarette and cigarette use changed over time?
```{r, echo = F, message=F}
ecig_data %>% 
  select(year, weight, ecigs_past_month, cig_days) %>% 
  #drop_na(ecigs_past_month) %>% 
  mutate(
    ecigs_past_month = recode(ecigs_past_month, 
                                     "0" = "0",
                                     "1-2" = "1.5",
                                     "3-5" = "4", 
                                     "6-9" = "7.5", 
                                     "10-19" = "14.5", 
                                     "20-29" = "24.5", 
                                     "30" = "30"),
    cig_days = recode(cig_days, 
                                     "0" = "0",
                                     "1-2" = "1.5",
                                     "3-5" = "4", 
                                     "6-9" = "7.5", 
                                     "10-19" = "14.5", 
                                     "20-29" = "24.5", 
                                     "30" = "30")
  ) %>% 
  group_by(year) %>% 
  summarize(
    cig_days = weighted.mean(as.numeric(cig_days), as.numeric(weight), na.rm = TRUE), 
    ecig_days = weighted.mean(as.numeric(ecigs_past_month), as.numeric(weight), na.rm = TRUE)
  ) %>%
  pivot_longer(cig_days:ecig_days, values_to = "mean_days", names_to = "type") %>% 
  filter(mean_days!="NaN") %>% 
  ggplot(aes(x = year, y = mean_days)) + 
  geom_line(aes(group = type, color = type)) + 
  geom_point(aes(color = type)) + 
  viridis::scale_color_viridis(
    discrete = TRUE, 
    name = "",
    breaks=c("cig_days", "ecig_days"),
    labels=c("Cigarettes", "E-Cigs")
    ) + 
  theme_bw() + 
  labs(title = "Average Days per Month Smoking", y = "Days Smoking", x = "Year") + 
  theme(legend.position = "bottom") 
  
```


Are cigarette smokers more likely to try electronic cigarettes?
```{r, echo = F, message=F}
#smokers are more likley to try ecigs
ecig_data %>%
  filter(is.na(age) == FALSE, is.na(try_ecigs) == FALSE, is.na(cig_days) == FALSE) %>%
  mutate(current_cig = as.numeric(cig_days != "0")) %>%
  mutate(current_cig = factor(current_cig, levels = c(0,1), labels = c("non-smoker", "smoker"))) %>%
  group_by(age, current_cig) %>%
  mutate(indc = ifelse(try_ecigs == "yes", 1, 0)) %>%
  summarise(y = weighted.mean(indc, as.numeric(weight))) %>%
  ggplot(aes(x = as.numeric(age), y = y, color = current_cig)) +
  geom_point() +
  geom_line() +
  xlab("Age") +
  ylab("Proportion Tried E-cigarettes") +
  labs(color = "Smoking Status") +
  theme_bw()
```



# Demographics

Does electronic cigarette use differ across age and sex?
```{r, echo = F, message=F}
ecig_data %>%
  filter(is.na(age) == FALSE, is.na(sex) == FALSE, is.na(try_ecigs) == FALSE) %>%
  group_by(age, sex) %>%
  mutate(indc = ifelse(try_ecigs == "yes", 1, 0)) %>%
  summarise(y = weighted.mean(indc, as.numeric(weight))) %>%
  ggplot(aes(x = age, y = y, fill = sex)) +
  geom_col(position = "dodge") +
  xlab("Age") +
  ylab("Proportion Tried E-cigarettes") +
  theme_minimal()
```



Which races use e-cigs the most? 
```{r}
ecig_data %>% 
  mutate(
    ecigs_past_month = as.numeric(ecigs_past_month), 
  ) %>% 
  select(ecigs_past_month, race, weight) %>% 
  mutate(race = fct_relevel(race, c("native hawaiian / other pacific islander", "american indian / alaska native", "white", "hispanic", "black", "asian"))) %>% 
  drop_na() %>% 
  group_by(race) %>% 
  summarize(ecig_use = round(weighted.mean(ecigs_past_month, as.numeric(weight)), digits = 3))  %>% 
  ggplot(aes(x = race, y = ecig_use)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  scale_x_discrete(name= "", labels = c("Native Hawaiian", "American Indian", "White", "Hispanic", "Black", "Asian")) + 
  labs(y = "Monthly E-Cig Use", title = "E-Cig Use across Races")
```


